---
title: 'Disaster Relief Project, Part 1'
author: "Becky Desrosiers, Abner Casillas-Colon, Rachel Daniel"
date: "2024-03-16"
output: pdf_document
---

```{r r-setup, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(cache = TRUE,
                      autodep = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      include = FALSE)
library(tidyverse)
library(tidymodels)
library(ggcorrplot)
library(GGally)
library(discrim)
library(patchwork)
library(doParallel)
cl <- makePSOCKcluster(parallel::detectCores(logical = FALSE))
registerDoParallel(cl)
```


# Introduction

The 2010 earthquake in Haiti was a devastating natural disaster that caused extreme damage and displaced millions of people. After this disaster, the rescue workers needed to deliver food, water, and other resources to the people of Haiti, but it was challenging locating the people in need over this large area. These challenges included roads blocked by rubble and the inability to communicate because of damage to communication infrastructure. The rescue workers needed to find another strategy other than physically looking for them on land to locate and reach these displaced persons quickly and more efficiently. 

The people of Haiti who were displaced by the earthquake were using blue tarps as temporary shelter. This knowledge was utilized to locate these people after imagery was collected by aircraft flown by a rescue team from the Rochester Institute of Technology. Blue tarps could be searched for within these images by the rescue team who would then go to these coordinates and find them. However, that strategy would have been too slow, and resources would not have been delivered in time. A different strategy that could be used to more efficiently locate people in need of resources was to continue to use these images but instead use data-mining algorithms to search these images. 

This report explores various classification methods that could be useful in locating these blue tarps within the images taken by aircraft. The algorithms tested and explored include those that utilize Logistic Regression with and without penalty, Linear Discriminant Analysis, Quadratic Discriminant Analysis, and K-Nearest Neighbors. We expect at least one of these methods to work more efficiently and more accurately than human efforts as these algorithms can handle more data, perform complex operations, and potentially fit to the patterns of the data to identify the blue tarps. The chosen algorithm will likely to miss fewer blue tarps than a human would. It is critical to identify that algorithm that will perform the best so that aid can be delivered to displaced people in dire circumstances. This is an important data-mining problem that could have a large impact on human life. This report will identify the algorithm that most accurately and most efficiently identifies these blue tarps so that the displaced Haitians can receive desperately needed resources in time. 


# Data

```{r load-data, warning = FALSE, message = FALSE}
# load data from file
# file = "https://gedeck.github.io/DS-6030/project/HaitiPixels.csv"
file = '../data.csv'
data <- read_csv(file) #%>% 
  #mutate(Class = factor(ifelse(Class == 'Blue Tarp', 'Tarp', 'Non-Tarp')))

data %>% glimpse()
data %>% summary()
```



The boxplots below showcase the distribution of the range of pixel values grouped by each of the categories of images observed. For are analysis one thing we would like to note is the potential groups that could overlap with out goal value of blue tarps. Rooftop values seem to be the closet for which a potential prediction error is likely to occur as its values overlap quite closely with the blue tarps. This is followed by Various Non-Tarps that have a large range of values which encompass tarps.

```{r EDA_Boxplots}
#| fig.width: 8
#| fig.height: 8
#| fig.align: center
#| out.width: 75%
g1_box <- ggplot(data, aes(x=Class, y=Red))+
  geom_boxplot()+
  labs(title= "Boxplots of Red Pixel Value by Class")

g2_box <- ggplot(data, aes(x=Class, y=Green))+
  geom_boxplot()+
  labs(title= "Boxplots of Green Pixel Value by Class")

g3_box <- ggplot(data, aes(x=Class, y=Blue))+
  geom_boxplot()+
  labs(title= "Boxplots of Blue Pixel Value by Class")

g1_box/g2_box/g3_box

```

The countplot below shows the total counts grouped by class. Blue tarp has the smallest amount of values at 2022. While this is important to note for the threshold selection later on in the model. It is promising that the two largest values by far are vegetation and soil as they are the two who are least likely to represent false positive or false negative values when predicting for tarps. Rooftops and Various Non-Tarps however, do account for 14,647 which do have closer overlaps.  

```{r EDA_Countplot_Class_ungrouped}

ggplot(data, aes(Class))+
  geom_bar()+
  labs(title= "Count Plot of Class", x="Count")

table(data$Class) %>% knitr::kable()
```



```{r EDA_Count_Table}
data <- data #%>% 
  mutate(Class = factor(ifelse(Class == 'Blue Tarp', 'Tarp', 'Non-Tarp')))
table(data$Class) %>% knitr::kable()
```

# Description of Methodology

This model is created using RStudio with the following packages: tidyverse, tidymodels, ggcorrplot, GGally, patchwork, discrim and doParallel. RStudio is an open-source statistical analysis software that offers a variety of packages to assist in model building and model selection. The first step of this analysis is to change the classifier into our variable of interest. The raw data classifies the pixels into Blue Tarp, Rooftop, Soil, Various Non-Tarp, and Vegetation, however we are only interested in whether the pixel is a Blue Tarp or not a Blue Tarp. We changed the factor outcome variable to have only two levels: Tarp and Non-Tarp. As the goal of this model will be to predict the class of a combination of pixels the following classification models will be explored: Logistic Regression, Penalized Logistic Regression (a logistic regression model with tuning parameters for an elastic net), linear discriminant analysis (LDA), quadtratic discriminant analysis (QDA), and a K-Nearest-Neighbors (KNN) model.

Of these models, the KNN model and the Logistic Regression Model will utilize tuning parameters. Tuning parameters will be selected to maximuze ROC AUC. For the KNN model, the selected tuning parameter is the number of neighbors, k. The number of neighbors selected will be the lowest k within one standard deviation of the maximum ROC AUC. For the penalized model, mixture and penalty will be tuned, and the best model will be selected based on ROC AUC.

For all models a tenfold cross-validation will be used for model validation. All metrics will be estimated using tenfold cross-validation. With an overall sample size of 63,241, this allows for each fold to have 6,324 records. With this amount of data, a tenfold cross validation is a reliable method to evaluate the performance of the training sets for all models.

All models will be compared based on ROC AUC. This metric provides a method to evaluate the performance of the True Positive Rate and False Positive Rate across all possible thresholds from 0 to 1. With its efficacy as an overall evaluator of performance, this metric will be used for model selection when comparing all versions of the tenfold validation models. 

As established during the EDA, the proportion of Blue Tarps to all other classes is approximately 3%. This indicates that a 50% threshold would be a poor choice for the final model. While five metrics will be explored for threshold selection (accuracy, j-index, sensitivity, specificity, and precision) our primary metrics of interest will be sensitivity and j-index. Sensitivity is especially important because we want to minimize false negatives - those could represent individuals that could be missed when response time is extremely important. The j-index is chosen as a metric that balances sensitivity and specificity. This will be valuable in models were responders have limited resources to check potential points of interest. While maximizing based on j-index will result in a higher false negative rate, we will greatly reduce false positives, which will make sure that we can make the most out of the time and resources allocated to the relief effort. In this report, we will assume that we have relatively unlimited resources and therefore pick thresholds based rimarily on sensitivity. J-index will be our secondary metric for selecting threshold. These same metrics will be used and weighted similarly when determining model performance, for similar reasons.



# Results of Model Fitting, Tuning Parameter Selection, and Evaluation


```{r convenience-functions}
# define functions to use later for convenience
cv_control <- control_resamples(save_pred = TRUE)

roc_plot <- function(model_preds) {
  model_preds %>% 
    roc_curve(truth = Class,
              .pred_Tarp,
              event_level = 'second') %>% 
    autoplot()
}

threshold_metric_plot <- function(thresh_perf, max_sens, max_j) {
  ggplot(thresh_perf, aes(x = .threshold, y = .estimate, color = .metric)) +
    geom_line() +
    geom_vline(data = max_sens, aes(xintercept = .threshold, color = .metric)) +
    geom_vline(data = max_j, aes(xintercept = .threshold, color = .metric)) +
    scale_x_continuous(breaks = seq(0, 1, 0.1)) +
    labs(x = 'Threshold', y = 'Metric value')
}

threshold_metrics <- metric_set(j_index,
                                sensitivity,
                                specificity,
                                accuracy,
                                precision)


threshold_preds <- function(preds, threshold) {
    preds <- preds %>% 
    mutate(.pred_class = factor(ifelse(.pred_Tarp >= threshold, 'Tarp', 'Non-Tarp')))
}

```


```{r model-setup}
# prepare resamples for 10-fold cross-validation
set.seed(6030)
resamples <- vfold_cv(data, v = 10, strata=Class)

# define formula
formula <- Class ~ Red + Green + Blue

# define basic recipe
rec <- recipe(formula, data = data)
```


## K-Nearest Neighbors


```{r knn}
# define model spec
knn_spec <- nearest_neighbor(mode = 'classification',
                            engine = 'kknn',
                            neighbors = parsnip::tune())

# define workflow
knn_wf <- workflow() %>% 
  add_recipe(rec) %>% 
  add_model(knn_spec)

# set tuning parameters
knn_params <- extract_parameter_set_dials(knn_wf) %>% 
  update(neighbors = neighbors(c(2, 100)))

# tune with grid (or fit resamples)
knn_tune <- tune_grid(knn_wf,
                      resamples = resamples,
                      grid = grid_regular(knn_params, levels = 50),
                      control = cv_control
                      )

# get tuning results visualization
knn_tune_vis <- autoplot(knn_tune, metric = 'roc_auc') +
  geom_hline(yintercept = (0.9943110 - 0.001139833), linetype = 'dashed')

# finalize workflow and fit resamples with best parameters
knn_best_params <- select_by_one_std_err(knn_tune, neighbors, metric = 'roc_auc')

knn_fitcv <- knn_wf %>% 
  finalize_workflow(knn_best_params) %>% 
  fit_resamples(resamples,
                control = cv_control
                )

# collect predictions and roc_auc from cross-validated fit
knn_preds <-  collect_predictions(knn_fitcv)
knn_metrics <- collect_metrics(knn_fitcv)

# get ROC plot
knn_roc <- roc_plot(knn_preds)

# get threshold selection info
knn_thresh_perf <- probably::threshold_perf(knn_preds,
                                            Class,
                                            .pred_Tarp,
                                            thresholds = seq(0.01, 0.99, 0.01),
                                            event_level = 'second',
                                            metrics = threshold_metrics)

# get threshold for best sensitivity
knn_max_sens <- knn_thresh_perf %>% 
  filter(.metric == 'sensitivity') %>% 
  filter(.estimate == max(.estimate))

# get threshold for best j-index
knn_max_j <- knn_thresh_perf %>% 
  filter(.metric == 'j_index') %>% 
  filter(.estimate == max(.estimate))

# get threshold metrics plot
knn_tmetrics_plot <- threshold_metric_plot(knn_thresh_perf, knn_max_sens, knn_max_j)

# get predictions and metrics based on chosen threshold
knn_threshold_preds <- knn_preds %>% 
  mutate(.pred_class = factor(ifelse(.pred_Tarp >= 0.04, 'Tarp', 'Non-Tarp')))

knn_threshold_metrics <- threshold_metrics(knn_threshold_preds,
                                           truth = Class,
                                           estimate = .pred_class)

# get confusion matrix for chosen threshold
knn_cm <- conf_mat(knn_threshold_preds, truth = Class, estimate = .pred_class)
```


We will now build a K-nearest neighbors (KNN) model with the number of neighbors, k, tuned to find the best fit. Figure X shows the ROC AUC of the model when fitted with different numbers of neighbors. It shows that the AUC of the ROC curve increases with k, maxing around 50, and leveling out at higher tuning values. The highest AUC is **0.9943** and belongs to the model with k = 40 neighbors. However, we choose a model within one standard deviation of the best metric for simplicity. From the plot, we can see that the point with the lowest number of neighbors above the dashed standard-error line represents the model with k = 8 nearest neighbors. We will use this tuning metric moving forward.


```{r figure-X-knn-tune, output = TRUE}
#| fig.width: 6
#| fig.height: 5
#| fig.align: center
#| out.width: 70%
#| fig.cap: ROC AUC based on number of principal components and nearest neighbors. The dashed horizontal line represents one standard deviation lower than the best AUC metric.
#| dev: "png"
#| dpi: 100
knn_tune_vis
```

With the number of neighbors tuned to k = 8, our model has an ROC AUC of **0.9933**. The ROC curve will be displayed in the next section, when all of the models are compared. The next step is to select the appropriate threshold for our model. Figure X shows five different metrics plotted against the threshold value, with thresholds ranging from 0.01 to 0.99. The metrics were calculated using tenfold cross-validation. The vertical lines show that sensitivity and j-index are both maximized at 0.02, 0.03, and 0.04 equally. Therefore, since the other metrics increase with the threshold, we choose the highest one, threshold = 0.04, to move forward with.

```{r figure-X-knn-threshold-metrics, output = TRUE}
#| fig.align: center
#| out.width: 70%
#| fig.cap: Five metrics based on threshold value for the KNN model with k = 8 nearest neighbors. Vertical lines show maximum j-index and sensitivity between 0.02 and 0.04.
#| dev: "png"
#| dpi: 100
knn_tmetrics_plot
```

Our final model has k = 8 nearest neighbors and a selection threshold of 0.04. The corresponding confusion matrix is shown below.

```{r knn-conf-mat, output = TRUE}
knn_cm
```

The confusion matrix shows that our final model misses only 26 Blue Tarp pixels and falsely identifies 251 Non-Tarp pixels. The overall accuracy of the model is 0.9956, with sensitivity of 0.9959. The precision is also very high: 0.9996. The specificity is 0.9871 and the j-index comes out to 0.9830.


# Linear Discriminant Analysis

```{r lda}
# define model spec
lda_spec <- discrim_linear(mode = "classification") %>% 
  set_engine('MASS')

# define workflow
lda_wf <- workflow() %>% 
  add_recipe(rec) %>% 
  add_model(lda_spec)


# fit resamples
lda_fitcv <- fit_resamples(lda_wf, resamples,  control=cv_control)

# collect predictions and roc_auc from cross-validated fit
lda_preds <-  collect_predictions(lda_fitcv)
lda_metrics <- collect_metrics(lda_fitcv)

lda_roc <- roc_plot(lda_preds)

# get threshold selection info
lda_thresh_perf <- probably::threshold_perf(lda_preds,
                                            Class,
                                            .pred_Tarp,
                                            thresholds = seq(0.01, 0.99, 0.01),
                                            event_level = 'second',
                                            metrics = metric_set(j_index,
                                                                 specificity,
                                                                 sensitivity,
                                                                 accuracy))

# get threshold for best sensitivity
lda_max_sens <- lda_thresh_perf %>% 
  filter(.metric == 'sensitivity') %>% 
  filter(.estimate == max(.estimate))

# get threshold for best j-index
lda_max_j <- lda_thresh_perf %>% 
  filter(.metric == 'j_index') %>% 
  filter(.estimate == max(.estimate))

# get threshold metrics plot
lda_tmetrics_plot <- threshold_metric_plot(lda_thresh_perf, lda_max_sens, lda_max_j)

# get predictions and metrics based on chosen threshold
lda_threshold_preds <- lda_preds %>% 
 mutate(.pred_class = factor(ifelse(.pred_Tarp >= 0.01, 'Tarp', 'Non-Tarp')))

lda_threshold_metrics <- threshold_metrics(lda_threshold_preds, truth = Class, estimate = .pred_class)


# get confusion matrix for chosen threshold
lda_cm <- conf_mat(lda_threshold_preds, truth = Class, estimate = .pred_class)

```

We built and trained the LDA model using the data set and then assessed the model’s performance metrics using 10-fold cross validation where we set the seed for reproducibility. We found that the model has an AUC of 0.9888, which means that it performs well on the training data and can classify blue tarp pixels versus non-tarp pixels. The next step was choosing the appropriate threshold for this model. Figure X displays the threshold plot with the plotted values of the chosen metrics from 0 to 1.  

```{r figure-X-lda-threshold-metrics, output = TRUE}
#| fig.align: center
#| out.width: 70%
#| fig.cap: Four metrics based on threshold value for the LDA model. Vertical lines show maximum j-index and sensitivity at 0.01.
#| dev: "png"
#| dpi: 100
lda_tmetrics_plot
```
 

These displayed metrics are accuracy, j-index, sensitivity, and specificity, but we are most interested in sensitivity and j-index. Both sensitivity and j-index perform the best at a threshold of around 0.01. This value of 0.01 will be the chosen threshold for our model when utilizing it for predictions. After threshold selection, the confusion matrix for the model was assessed where we explored the number of false negatives, false positives, and True positives (figure X). We found that while we miss some blue tarp pixels, we locate many more, while also limiting the number of false positives.
 

```{r lda-cm, output = TRUE}
lda_cm
```

 
The LDA model performed very well on the data at the selected threshold. Sensitivity, accuracy, and precision are all estimated between 0.976 and 0.996, which is very high for these metrics. This model appears to be accurate and sensitive to this data. The LDA model could be a good option as our chosen model, but it will need to be compared to the other models first, which we will do later in this report. 


# Quadratic Discriminant Analysis

```{r qda}
# define model spec
qda_spec <- discrim_quad(mode = "classification") %>% 
  set_engine('MASS')

# define workflow
qda_wf <- workflow() %>% 
  add_recipe(rec) %>% 
  add_model(qda_spec)


#fit resamples
qda_fitcv <- fit_resamples(qda_wf, resamples,  control=cv_control)

# collect predictions and roc_auc from cross-validated fit
qda_preds <-  collect_predictions(qda_fitcv)
qda_metrics <- collect_metrics(qda_fitcv)

qda_roc <- roc_plot(qda_preds)

# get threshold selection info
qda_thresh_perf <- probably::threshold_perf(qda_preds,
                                            Class,
                                            .pred_Tarp,
                                            thresholds = seq(0.01, 0.99, 0.01),
                                            event_level = 'second',
                                            metrics = metric_set(j_index,
                                                                 specificity,
                                                                 sensitivity,
                                                                 accuracy))

# get threshold for best sensitivity
qda_max_sens <- qda_thresh_perf %>% 
  filter(.metric == 'sensitivity') %>% 
  filter(.estimate == max(.estimate))

# get threshold for best j-index
qda_max_j <- qda_thresh_perf %>% 
  filter(.metric == 'j_index') %>% 
  filter(.estimate == max(.estimate))

# get threshold metrics plot
qda_tmetrics_plot <- threshold_metric_plot(qda_thresh_perf, qda_max_sens, qda_max_j)


# get predictions and metrics based on chosen threshold
qda_threshold_preds <- qda_preds %>% 
 mutate(.pred_class = factor(ifelse(.pred_Tarp >= 0.015, 'Tarp', 'Non-Tarp')))

qda_threshold_metrics <- threshold_metrics(qda_threshold_preds, truth = Class, estimate = .pred_class)


# get confusion matrix for chosen threshold
qda_cm <- conf_mat(qda_threshold_preds, truth = Class, estimate = .pred_class)
```


We built and trained a QDA model using blue tarp and non-tarp data. This model will potentially be used to predict blue tarp locations so that resources can be provided to people in need. We found that the model's AUC was 0.9982, which means it performs very well on the data. The QDA model can almost perfectly classify the data set into blue tarp pixels versus non-tarp pixels. The next step was selecting the best threshold based on sensitivity and j_index. We found that the maximum sensitivity is at a threshold of 0.01, while the maximum j-index was at 0.02. The metrics are revealed in the threshold plot in figure X. 

```{r figure-X-qda-threshold-metrics, output = TRUE}
#| fig.align: center
#| out.width: 70%
#| fig.cap: Four metrics based on threshold value for the QDA model. Vertical lines show maximum j-index and sensitivity at 0.02 and 0.01, respectively.
#| dev: "png"
#| dpi: 100
qda_tmetrics_plot
```
 

Since sensitivity and j-index are both important for this data, we decided to choose a threshold in between the two for our model, which was 0.015. This threshold value seems to maximize these metrics, as the j-index comes out to be 0.9692, while the sensitivity is 0.9717.  

After we selected the appropriate threshold for our mode, we then assessed the confusion matrix for this model with the selected threshold. The model performed extremely well at this selected threshold in terms of locating blue tarp pixels as there are only 5 that are missed. However, when choosing a final model, it will be important to keep in mind that the number of false positives is almost equal to the number of true positives for this model. 


```{r qda-cm, output = TRUE}
qda_cm
```

The QDA model at the selected threshold has a sensitivity of 0.9717, accuracy of 0.9725, and precision of 0.9999. These metrics reveal that the QDA model performs very well on the data and can correctly identify almost all the blue tarp pixels. The QDA model appears to perform better than the LDA model so it seems that the data benefits from some flexibility in the predictive algorithm. 



# Logistic Regression Analysis

```{r logistic_regression}
log_spec <- logistic_reg(mode = 'classification',
                            engine = 'glm')
                            
# define workflow
log_wf <- workflow() %>% 
  add_recipe(rec) %>% 
  add_model(log_spec)


# tune with grid (or fit resamples)

log_fitcv <-fit_resamples(log_wf, resamples, control=cv_control)

# collect predictions and roc_auc from cross-validated fit
log_preds <-  collect_predictions(log_fitcv)
log_metrics <- collect_metrics(log_fitcv)

# get ROC plot
log_roc <- roc_plot(log_preds)

# get threshold selection info
log_thresh_perf <- probably::threshold_perf(log_preds,
                                            Class,
                                            .pred_Tarp,
                                            thresholds = seq(0.01, 0.99, 0.01),
                                            event_level = 'second',
                                            metrics = metric_set(j_index,
                                                                 specificity,
                                                                 sensitivity,
                                                                 accuracy))

# get threshold for best sensitivity
log_max_sens <- log_thresh_perf %>% 
  filter(.metric == 'sensitivity') %>% 
  filter(.estimate == max(.estimate))

# get threshold for best j-index
log_max_j <- log_thresh_perf %>% 
  filter(.metric == 'j_index') %>% 
  filter(.estimate == max(.estimate))

# get threshold metrics plot
log_tmetrics_plot <- threshold_metric_plot(log_thresh_perf, log_max_sens, log_max_j)

# get predictions and metrics based on chosen threshold
log_threshold_preds <- log_preds %>% 
  mutate(.pred_class = factor(ifelse(.pred_Tarp >= 0.04, 'Tarp', 'Non-Tarp')))

log_threshold_metrics <- threshold_metrics(log_threshold_preds,
                                           truth = Class,
                                           estimate = .pred_class)

# get confusion matrix for chosen threshold
log_cm <- conf_mat(log_threshold_preds, truth = Class, estimate = .pred_class)
```

# Penalized Logistic Regression

```{r penalized_logistic}
# define model spec
pen_spec <- logistic_reg(mode = 'classification',
                            engine = 'glmnet',
                            penalty = tune(),
                            mixture = tune())
                            
# define workflow
pen_wf <- workflow() %>% 
  add_recipe(rec) %>% 
  add_model(pen_spec)

# set tuning parameters
pen_params <- extract_parameter_set_dials(pen_wf) %>% 
  update(
    penalty=penalty(c(-20,-1)),
    mixture=mixture(c(0,1))
  )

# tune with grid (or fit resamples)
pen_tune <- tune_grid(pen_wf,
                      resamples=resamples,
                      control=cv_control,
                      grid=grid_latin_hypercube(pen_params, size=50))

# get tuning results visualization
pen_tune_vis <- autoplot(pen_tune)$roc_auc

# finalize workflow and fit resamples with best parameters
pen_best_params <- select_best(pen_tune, metric="roc_auc")

pen_fitcv <- pen_wf %>% 
  finalize_workflow(pen_best_params) %>% 
  fit_resamples(resamples,
                control = cv_control
                )

# collect predictions and roc_auc from cross-validated fit
pen_preds <-  collect_predictions(pen_fitcv)
pen_metrics <- collect_metrics(pen_fitcv)

# get ROC plot
pen_roc <- roc_plot(pen_preds)

# get threshold selection info
pen_thresh_perf <- probably::threshold_perf(pen_preds,
                                            Class,
                                            .pred_Tarp,
                                            thresholds = seq(0.01, 0.99, 0.01),
                                            event_level = 'second',
                                            metrics = metric_set(j_index,
                                                                 specificity,
                                                                 sensitivity,
                                                                 accuracy))

# get threshold for best sensitivity
pen_max_sens <- pen_thresh_perf %>% 
  filter(.metric == 'sensitivity') %>% 
  filter(.estimate == max(.estimate))

# get threshold for best j-index
pen_max_j <- pen_thresh_perf %>% 
  filter(.metric == 'j_index') %>% 
  filter(.estimate == max(.estimate))

# get threshold metrics plot
pen_tmetrics_plot <- threshold_metric_plot(pen_thresh_perf, pen_max_sens, pen_max_j)

# get predictions and metrics based on chosen threshold
# the optimal threshold for the sens is .01
pen_threshold_preds <- pen_preds %>% 
  mutate(.pred_class = factor(ifelse(.pred_Tarp >= 0.04, 'Tarp', 'Non-Tarp')))

penlog_threshold_metrics <- threshold_metrics(pen_threshold_preds,
                                           truth = Class,
                                           estimate = .pred_class)

# get confusion matrix for chosen threshold
pen_cm <- conf_mat(pen_threshold_preds, truth = Class, estimate = .pred_class)
```



```{r compare-models, output = TRUE}
aucs <- ( log_roc + penlog_roc  + knn_roc) / ( lda_roc + qda_roc )

compare_metrics <- bind_rows(log_threshold_metrics %>% mutate(model = 'logistic regression'),
                             penlog_threshold_metrics %>% mutate(model = 'penalized logreg'),
                             lda_threshold_metrics %>% mutate(model = 'LDA'),
                             qda_threshold_metrics %>% mutate(model = 'QDA'),
                             knn_threshold_metrics %>% mutate(model = 'KNN')
                             ) %>% 
  pivot_wider(id_cols = model,
              names_from = .metric,
              values_from = .estimate)
```





# Conclusion


>Three or more clearly identifiable conclusions. This section is more important than the previous sections (as
reflected in the points). Give sufficient explanation and justification for each conclusion.
One conclusion must be:
• determination and justification of which algorithm works best.
Additional conclusions should be observations you’ve made based on your work on this project, such as:
• What additional recommend actions can be taken to improve results?
• Were there multiple adequately performing methods, or just one clear best method? What is your level
of confidence in the results?
• What is it about this data formulation that allows us to address it with predictive modeling tools?
• How effective do you think your work here could actually be in terms of helping to save human life?
• Do these data seem particularly well-suited to one class of prediction methods, and if so, why?
These are only suggestions, pursue your own interests. Your best two additional conclusions will be graded.
Make sure that the 3 conclusions are clearly separated.

## Conclusion 1

We concluded that the KNN method results in the best algorithm that classifies each pixel as a blue tarp pixel or a non-tarp pixel after the performance metrics for each model were assessed. While it is important to consider all metrics when determining the best model, in this context, it is most important to identify the majority of the tarp pixels and limit the number of false negatives so that we can locate most of the displaced persons. The sensitivity and j-index were the metrics that we mainly focused on with this goal in mind. The KNN model had the highest j-index and the highest sensitivity out of all of the models so that is the one that we believe will be most useful and accurate in identifying blue tarps. This model will play a major role in providing resources to displaced persons and potentially saving human lives.


## Conclusion 3

We found that all of the methods performed very well, and there was no one clearly best method. Though the KNN model had the best sensitivity and j-index, which we considered to be the most important metrics, all of the models had very high metrics for not only j-index and sensitivity, but specificity, accuracy and precision as well. For this reason, we are more confident in our results because the models were able to reliably perform well on the data.






# Appendix

```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
stopCluster(cl)
registerDoSEQ()
```
